---
output: html_document
---
Characteristics of new hires
========================================================
We've been analysing new hires (people starting jobs in the previous 3 months) by occupation and qualification (and region)as part of our skills work.

As we've worked out how to generate datasets of new hires, we can cut them any sensible way (and some not so sensible). 

Using 8 quarter pooled datasets of new hires at UK level, we have done occupation (3-digit at region, 4-digit at national), qualification level.

If we have equivalent pooled datasets for pre-recession, recession and post-recession periods, we can illustrate changes in absolute numbers of new hires, and changes in the breakdowns by:

* age
* gender
* ethnicity
* disability
* contract type
* as well as region, qualification and occupation
* Not sure I want to go into sector - although I trust LFS sectors more than employer-based (the opposite of ONS)

N.B. New hires are weighted to high-turnover occupations, and do include people moving from previous jobs.

### Dataset preparation

Use four time periods where available, 1996-97, 2009-7, 2009-10, 2012-13 so can use bar charts. Alternatively line charts with superimposed rectangles (but bars are easier).

```{r required_packages, cache=TRUE, echo=FALSE}
require(memisc)
require(survey)
```

```{r load_data, cache=TRUE, echo=FALSE}
list_files <- list.files("K:/data/LFS/Quarterly", pattern=glob2rx("qlfs*.sav"), full.names=TRUE)
#1996-97
aj96 <- spss.system.file(list_files[19])
aj97 <- spss.system.file(list_files[20])
jm96 <- spss.system.file(list_files[40])
jm97 <- spss.system.file(list_files[41])
js96 <- spss.system.file(list_files[61])
js97 <- spss.system.file(list_files[62])
od96 <- spss.system.file(list_files[82])
od97 <- spss.system.file(list_files[83])
# Subset data
aj96.data <- subset(aj96, select=c(emplen, age, sex, ethcen, govtor, hiquald, socmajm, inecacr, ftpt, jobtyp, jobtmp, pwt07))
aj97.data <- subset(aj97, select=c(emplen, age, sex, ethcen, govtor, hiquald, socmajm, inecacr, ftpt, jobtyp, jobtmp, pwt07))
jm96.data <- subset(jm96, select=c(emplen, age, sex, ethcen, govtor, hiquald, socmajm, inecacr, ftpt, jobtyp, jobtmp, pwt07))
jm97.data <- subset(jm97, select=c(emplen, age, sex, ethcen, govtor, hiquald, socmajm, inecacr, ftpt, jobtyp, jobtmp, pwt07))
js96.data <- subset(js96, select=c(emplen, age, sex, ethcen, govtor, hiquald, socmajm, inecacr, ftpt, jobtyp, jobtmp, pwt07))
js97.data <- subset(js97, select=c(emplen, age, sex, ethcen, govtor, hiquald, socmajm, inecacr, ftpt, jobtyp, jobtmp, pwt07))
od96.data <- subset(od96, select=c(emplen, age, sex, ethcen, govtor, hiquald, socmajm, inecacr, ftpt, jobtyp, jobtmp, pwt07))
od97.data <- subset(od97, select=c(emplen, age, sex, ethcen, govtor, hiquald, socmajm, inecacr, ftpt, jobtyp, jobtmp, pwt07))

measurement(aj96.data$age) <- "interval"
measurement(aj97.data$age) <- "interval"
measurement(jm96.data$age) <- "interval"
measurement(jm97.data$age) <- "interval"
measurement(js96.data$age) <- "interval"
measurement(js97.data$age) <- "interval"
measurement(od96.data$age) <- "interval"
measurement(od97.data$age) <- "interval"

measurement(aj96.data$pwt07) <- "interval"
measurement(aj97.data$pwt07) <- "interval"
measurement(jm96.data$pwt07) <- "interval"
measurement(jm97.data$pwt07) <- "interval"
measurement(js96.data$pwt07) <- "interval"
measurement(js97.data$pwt07) <- "interval"
measurement(od96.data$pwt07) <- "interval"
measurement(od97.data$pwt07) <- "interval"

pooled9697 <- rbind(as.data.frame(aj96.data),
                    as.data.frame(aj97.data),
                    as.data.frame(jm96.data),
                    as.data.frame(jm97.data),
                    as.data.frame(js96.data),
                    as.data.frame(js97.data),
                    as.data.frame(od96.data),
                    as.data.frame(od97.data))
pooled969b <- subset(pooled9697, emplen == "Less than 3 months")
pooled969b <- subset(pooled969b, age < 65)

#2006-07
aj06 <- spss.system.file(list_files[7])
aj07 <- spss.system.file(list_files[8])
jm06 <- spss.system.file(list_files[29])
jm07 <- spss.system.file(list_files[30])
js06 <- spss.system.file(list_files[50])
js07 <- spss.system.file(list_files[51])
od06 <- spss.system.file(list_files[71])
od07 <- spss.system.file(list_files[72])
# Subset data
aj06.data <- subset(aj06, select=c(emplen, age, sex, ethcen6, govtor, hiqual5d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07, discurr))
aj07.data <- subset(aj07, select=c(emplen, age, sex, ethcen6, govtor, hiqual5d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07 = PWT09, discurr))
jm06.data <- subset(jm06, select=c(emplen, age, sex, ethcen6, govtor, hiqual5d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07, discurr))
jm07.data <- subset(jm07, select=c(emplen, age, sex, ethcen6, govtor, hiqual5d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07 = PWT09, discurr))
js06.data <- subset(js06, select=c(emplen, age, sex, ethcen6, govtor, hiqual5d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07 = PWT09, discurr))
js07.data <- subset(js07, select=c(emplen, age, sex, ethcen6, govtor, hiqual5d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07 = PWT09, discurr))
od06.data <- subset(od06, select=c(emplen, age, sex, ethcen6, govtor, hiqual5d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07 = PWT09, discurr))
od07.data <- subset(od07, select=c(emplen, age, sex, ethcen6, govtor, hiqual5d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07 = PWT09, discurr))

measurement(aj06.data$age) <- "interval"
measurement(aj07.data$age) <- "interval"
measurement(jm06.data$age) <- "interval"
measurement(jm07.data$age) <- "interval"
measurement(js06.data$age) <- "interval"
measurement(js07.data$age) <- "interval"
measurement(od06.data$age) <- "interval"
measurement(od07.data$age) <- "interval"

measurement(aj06.data$pwt07) <- "interval"
measurement(aj07.data$pwt07) <- "interval"
measurement(jm06.data$pwt07) <- "interval"
measurement(jm07.data$pwt07) <- "interval"
measurement(js06.data$pwt07) <- "interval"
measurement(js07.data$pwt07) <- "interval"
measurement(od06.data$pwt07) <- "interval"
measurement(od07.data$pwt07) <- "interval"

pooled0607 <- rbind(as.data.frame(aj06.data),
                    as.data.frame(aj07.data),
                    as.data.frame(jm06.data),
                    as.data.frame(jm07.data),
                    as.data.frame(js06.data),
                    as.data.frame(js07.data),
                    as.data.frame(od06.data),
                    as.data.frame(od07.data))
pooled069b <- subset(pooled0607, emplen == "Less than 3 months")
pooled069b <- subset(pooled069b, age < 65)
# 2009-10
aj09 <- spss.system.file(list_files[10])
aj10 <- spss.system.file(list_files[11])
jm09 <- spss.system.file(list_files[32])
jm10 <- spss.system.file(list_files[33])
js09 <- spss.system.file(list_files[53])
js10 <- spss.system.file(list_files[54])
od09 <- spss.system.file(list_files[74])
od10 <- spss.system.file(list_files[75])
# Subset data
aj09.data <- subset(aj09, select=c(emplen, age, sex, ethcen6, govtor, hiqual8d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07=PWT09, discurr))
aj10.data <- subset(aj10, select=c(emplen, age, sex, ethcen6, govtor, hiqual8d, sc2kmmj, inecac05, ftpt, jobtyp, jbtp10, pwt07 = pwt09, discurr))
jm09.data <- subset(jm09, select=c(emplen, age, sex, ethcen6, govtor, hiqual8d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07=PWT09, discurr))
jm10.data <- subset(jm10, select=c(emplen, age, sex, ethcen6, govtor, hiqual8d, sc2kmmj, inecac05, ftpt, jobtyp, jbtp10, pwt07 = pwt09, discurr))
js09.data <- subset(js09, select=c(emplen, age, sex, ethcen6, govtor, hiqual8d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07 = PWT09, discurr))
js10.data <- subset(js10, select=c(emplen, age, sex, ethcen6, govtor, hiqual8d, sc2kmmj, inecac05, ftpt, jobtyp, jbtp10, pwt07 = pwt09, discurr))
od09.data <- subset(od09, select=c(emplen, age, sex, ethcen6, govtor, hiqual8d, sc2kmmj, inecac05, ftpt, jobtyp, jobtmp, pwt07 = pwt09, discurr))
od10.data <- subset(od10, select=c(emplen, age, sex, ethcen6, govtor, hiqual8d, sc2kmmj, inecac05, ftpt, jobtyp, jbtp10, pwt07 = pwt09, discurr))

measurement(aj09.data$age) <- "interval"
measurement(aj10.data$age) <- "interval"
measurement(jm09.data$age) <- "interval"
measurement(jm10.data$age) <- "interval"
measurement(js09.data$age) <- "interval"
measurement(js10.data$age) <- "interval"
measurement(od09.data$age) <- "interval"
measurement(od10.data$age) <- "interval"

measurement(aj09.data$pwt07) <- "interval"
measurement(aj10.data$pwt07) <- "interval"
measurement(jm09.data$pwt07) <- "interval"
measurement(jm10.data$pwt07) <- "interval"
measurement(js09.data$pwt07) <- "interval"
measurement(js10.data$pwt07) <- "interval"
measurement(od09.data$pwt07) <- "interval"
measurement(od10.data$pwt07) <- "interval"

pooled09 <- rbind(as.data.frame(aj09.data),
                    as.data.frame(jm09.data),
                    as.data.frame(js09.data),
                    as.data.frame(od09.data))
pooled10 <- rbind(as.data.frame(aj10.data),
                    as.data.frame(jm10.data),
                    as.data.frame(js10.data),
                    as.data.frame(od10.data))
pooled10$jobtmp <- recode(pooled10$jbtp10, 
                          "Seasonal work" <- "seasonal work",
                          "Contract for fixed period, fixed task" <- "contract for fixed period or fixed task",
                          "Agency temping" <- "working for employment agency",
                          "Casual work" <- "casual type of work", 
                          "Not permanent in some other way" <- "some other way")
pooled10 <- pooled10[,c(1:10,12:14)]
pooled0910 <- rbind(pooled09,pooled10)
pooled099b <- subset(pooled0910, emplen == "Less than 3 months")
pooled099b <- subset(pooled099b, age < 65)

#2011-13

aj12 <- spss.system.file(list_files[13])
aj13 <- spss.system.file(list_files[14])
#aj11 <- spss.system.file(list_files[12])
jm12 <- spss.system.file(list_files[35])
jm13 <- spss.system.file(list_files[36])
js12 <- spss.system.file(list_files[56])
js13 <- spss.system.file(list_files[57])
od12 <- spss.system.file(list_files[78])
od13 <- spss.system.file(list_files[79])
# Subset data
aj12.data <- subset(aj12, select=c(emplen = EMPLEN, age= AGE, sex= SEX , ethukeul = ETHUKEUL , govtor = GOVTOR, hiqul11d = HIQUL11D, sc10mmj = SC10MMJ , inecac05 = INECAC05, ftpt = FTPT , jobtyp = JOBTYP, jbtp101 = JBTP101, pwt07=PWT11, discurr = DISCURR))
aj13.data <- subset(aj13, select=c(emplen = EMPLEN, age= AGE, sex= SEX , ethukeul = ETHUKEUL , govtor = GOVTOR, hiqul11d = HIQUL11D, sc10mmj = SC10MMJ , inecac05 = INECAC05, ftpt = FTPT , jobtyp = JOBTYP, jbtp101 = JBTP101, pwt07=PWT11, discurr = DISCURR))
#aj11.data <- subset(aj11, select=c(emplen, age, sex, ethukeul, govtor, hiqul11d, sc10mmj, inecac05, ftpt, jobtyp, jbtp101, pwt07=pwt10, discurr))
jm13.data <- subset(jm13, select=c(emplen = EMPLEN, age= AGE, sex= SEX , ethukeul = ETHUKEUL , govtor = GOVTOR, hiqul11d = HIQUL11D, sc10mmj = SC10MMJ , inecac05 = INECAC05, ftpt = FTPT , jobtyp = JOBTYP, jbtp101 = JBTP101, pwt07=PWT11, discurr = DISCURR))
js12.data <- subset(js12, select=c(emplen = EMPLEN, age= AGE, sex= SEX , ethukeul = ETHUKEUL , govtor = GOVTOR, hiqul11d = HIQUL11D, sc10mmj = SC10MMJ , inecac05 = INECAC05, ftpt = FTPT , jobtyp = JOBTYP, jbtp101 = JBTP101, pwt07=PWT11, discurr = DISCURR))
#js11.data <- subset(js11, select=c(emplen, age, sex, ethukeul, govtor, hiqul11d, sc10mmj, inecac05, ftpt, jobtyp, jbtp101, pwt07=pwt10, discurr))
od12.data <- subset(od12, select=c(emplen = EMPLEN, age= AGE, sex= SEX , ethukeul = ETHUKEUL , govtor = GOVTOR, hiqul11d = HIQUL11D, sc10mmj = SC10MMJ , inecac05 = INECAC05, ftpt = FTPT , jobtyp = JOBTYP, jbtp101 = JBTP101, pwt07=PWT11, discurr = DISCURR))
#od11.data <- subset(od11, select=c(emplen, age, sex, ethukeul, govtor, hiqul11d, sc10mmj, inecac05, ftpt, jobtyp, jbtp101, pwt07=pwt10, discurr))
js13.data <- subset(js13, select=c(emplen = EMPLEN, age= AGE, sex= SEX , ethukeul = ETHUKEUL , govtor = GOVTOR, hiqul11d = HIQUL11D, sc10mmj = SC10MMJ , inecac05 = INECAC05, ftpt = FTPT , jobtyp = JOBTYP, jbtp101 = JBTP101, pwt07=PWT11, discurr = DISCURR13))
od13.data <- subset(od13, select=c(emplen = EMPLEN, age= AGE, sex= SEX , ethukeul = ETHUKEUL , govtor = GOVTOR, hiqul11d = HIQUL11D, sc10mmj = SC10MMJ , inecac05 = INECAC05, ftpt = FTPT , jobtyp = JOBTYP, jbtp101 = JBTP101, pwt07=PWT11, discurr = DISCURR13))
jm12.data <- subset(jm12, select=c(emplen = emplen, age= age, sex= sex , ethukeul = ethukeul , govtor = govtor, hiqul11d = hiqul11d, sc10mmj = sc10mmj , inecac05 = INECAC05, ftpt = ftpt , jobtyp = jobtyp, jbtp101 = jbtp101, pwt07=PWT11, discurr = discurr))



measurement(aj12.data$age) <- "interval"
measurement(aj13.data$age) <- "interval"
#measurement(aj11.data$age) <- "interval"
measurement(jm13.data$age) <- "interval"
measurement(jm12.data$age) <- "interval"
measurement(js12.data$age) <- "interval"
measurement(js13.data$age) <- "interval"
measurement(od12.data$age) <- "interval"
measurement(od13.data$age) <- "interval"

measurement(aj12.data$pwt07) <- "interval"
measurement(aj13.data$pwt07) <- "interval"
#measurement(aj11.data$pwt07) <- "interval"
measurement(jm12.data$pwt07) <- "interval"
measurement(jm13.data$pwt07) <- "interval"
measurement(js12.data$pwt07) <- "interval"
measurement(js13.data$pwt07) <- "interval"
measurement(od12.data$pwt07) <- "interval"
measurement(od13.data$pwt07) <- "interval"

pooled12 <- rbind(as.data.frame(jm12.data),
                    as.data.frame(aj12.data),
                    as.data.frame(js12.data),
                    as.data.frame(od12.data), 
                  as.data.frame(aj13.data),
                    as.data.frame(jm13.data),
                    as.data.frame(js13.data),
                    as.data.frame(od13.data))
pooled12$jobtmp <- recode(pooled12$jbtp101, 
                          "Seasonal work" <- c("seasonal work", "Seasonal work"),
                          "Contract for fixed period, fixed task" <- c("contract for fixed period or fixed task", "Contract for fixed period or fixed task", "Done under contract for a fixed period or for a fixed task"),
                          "Agency temping" <- c("working for employment agency", "Working for an employment agency", "Working for employment agency"),
                          "Casual work" <- c("casual type of work", "Casual type of work"), 
                          "Not permanent in some other way" <- c("some other way","Some other way", "Or was there some other way that it was not permanent" ))
pooled1213 <- pooled12[,c(1:10,12:14)]
pooled129b <- subset(pooled1213, emplen == "Less than 3 months")
pooled129b <- subset(pooled129b, age < 65)

pooled129b$sc10mmj <- recode(pooled129b$sc10mmj, 
     "Managers, Directors and Senior Officials" <- 
          c("Managers, Directors and Senior Officials", 
            "1  'Managers, Directors And Senior Officials'", 
            "Managers, Directors And Senior Officials"), 
     "Professional Occupations" <- 
          c("Professional Occupations", 
            "2  'Professional Occupations'"),
     "Associate Professional and Technical Occupations" <- 
          c("Associate Professional and Technical Occupations", 
            "3  'Associate Professional And Technical Occupations", 
            "Associate Professional And Technical Occupations"),
     "Administrative and Secretarial Occupations " <- 
       c("Administrative and Secretarial Occupations", 
         "4  'Administrative And Secretarial Occupations'", 
         "Administrative And Secretarial Occupations"),
      "Skilled Trades Occupations" <- 
       c("Skilled Trades Occupations", 
         "5  'Skilled Trades Occupations'"), 
                           "Caring, Leisure and Other Service Occupations" <- c("Caring, Leisure and Other Service Occupations", "6  'Caring, Leisure And Other Service Occupations'", "Caring, Leisure And Other Service Occupations"), 
                           "Sales and Customer Service Occupations" <- c("Sales and Customer Service Occupations","7  'Sales And Customer Service Occupations'", "Sales And Customer Service Occupations"), 
                           "Process, Plant and Machine Operatives" <- c("Process, Plant and Machine Operatives", "8  'Process, Plant And Machine Operatives'", "Process, Plant And Machine Operatives"), 
                           "Elementary Occupations" <- c("Elementary Occupations", "9  'Elementary Occupations'")
                           )
pooled129b$hiqul11d <- recode(pooled129b$hiqul11d, 
                              "GCSE grades A*-C or equivalent" <- c("GCSE grades A*-C or equivalent", "GCSE grades A-C or equiv"), otherwise="copy")
  
pooled96 <- svydesign(~1, data=pooled969b, weights=pooled969b$pwt07)
pooled06 <- svydesign(~1, data=pooled069b, weights=pooled069b$pwt07)
pooled09 <- svydesign(~1, data=pooled099b, weights=pooled099b$pwt07)
pooled12svy <- svydesign(~1, data=pooled129b, weights=pooled129b$pwt07)
pooled12svy_25 <- subset(pooled12svy, age >24)

```