---
title: "Productivity"
output: 
  flexdashboard::flex_dashboard
    
---

```{r setup, include=FALSE}
library(flexdashboard)
library(jsonlite)
library(dplyr)
library(tidyr)
library(xts)
library(forecast)
library(dygraphs)
library(metricsgraphics)
library(OECD)
library(forcats)
lzvb <- fromJSON("http://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/labourproductivity/timeseries/lzvb/prdy/data")
lzvb <- zooreg(as.numeric(lzvb$quarter$value), start=c(1971,1), frequency = 4)
OECD_PDB_LV <- get_dataset("PDB_LV", filter = "/CAN+FRA+DEU+ITA+JPN+GBR+USA+EA19+EU28+G-7+OECD.T_GDPPOP+T_GDPHRS.CPC",start_time = 2013, end_time = 2015, pre_formatted = TRUE)
OECD_PDB_LV2 <- OECD_PDB_LV[OECD_PDB_LV$SUBJECT == "T_GDPHRS",]
OECD_PDB_LV2 <- OECD_PDB_LV2[OECD_PDB_LV2$obsTime == "2014",]
OECD_PDB_LV2$LOCATION <- fct_reorder(OECD_PDB_LV2$LOCATION, OECD_PDB_LV2$obsValue)


```

Productivity {.storyboard}
==============================================================

### Productivity in the UK. Productivity is measured here as output (GDP) per hour worked, and so takes account of the effect of changes in working hours. Prices are held constant

```{r}
dygraph(as.xts(lzvb, as.yearqtr(index(lzvb)))) %>% 
  dyRangeSelector(dateWindow = c("2006-03-01", "2016-06-01"))

```

*** 

Before the recession, productivity in the UK was rising steadily. Since the recession, output per hour worked has not reurned to the same level that it was in the fourth quarter or 2007, let alone returned to steady growth.

Productivity in the UK has recovered from the immediate effects of the 2008 recession, but suffered a further decline from 2011 to 2013. Only in the latest yeas has productivity returned to 2007 levels.

If productivity had returned to it's pre-2007 growth pattern, it would be very much higher than it is in the latest figures.

### International comparisons of Productivity - The UK has the second lowest productivity per hour in the G7 group of developed countries (after Japan)

```{r}
OECD_PDB_LV2 %>%
  mjs_plot(y=LOCATION, x=obsValue, width=500, height=400) %>%
  mjs_bar() %>%
  mjs_axis_x(xax_format = 'plain')

```

*** 

The UK's productivity per hour worked was $`r I(round(OECD_PDB_LV2[OECD_PDB_LV2$LOCATION == "GBR","obsValue"],2))` in 2014. This was lower than that for the G7 countries as a whole (at $`r I(round(OECD_PDB_LV2[OECD_PDB_LV2$LOCATION == "G-7","obsValue"],2))`), for the EU as a whole ($`r I(round(OECD_PDB_LV2[OECD_PDB_LV2$LOCATION == "EU28","obsValue"],2))`) and the Euro Area ($`r I(round(OECD_PDB_LV2[OECD_PDB_LV2$LOCATION == "EA19","obsValue"],2))`). 

Among the G7 countries, the UK's productivity was only higher than that of Japan.

The productivity figures are measured in US dollars, adjusted by the OECD to 'purchasing power parity'. Figures are sourced from the OECD.

### Forecasts for UK productivity

```{r}

```

***

Some commentary about forecasts for UK Productivity
